\section{\texorpdfstring{Random number generators}{Random number generators}}
\vspace{5mm}
\large

\begin{definition}
	Secure RNG - even knowing complete past output, next bit is unpredictable $\Rightarrow$ statistically uniform.
\end{definition}

We have 2 type of random number generators:\\
a) PRNG: e.g. block cipher in CTR mode. Key should be changed on regular basis.\\
b) physical randomness (we have no proof such thing exists).
Examples:

\begin{itemize}
	\item Thermal noise or resister/diode
	\item Radioactive decay (count \# of particles)
	\item Radio noise
	\item Lava lamp (chaotic movement)
	\item Ring oscillator (attacker can observer)
	\item Precise timings of keystrokes, mouse movements or network packages (attacker can record or flood packages)
\end{itemize}

\begin{note}
	Entropy in physical processes is not as high as we can expect.
	Therefore, a combination of PRNG and physical randomness is the best choice.
\end{note}

Problem with PRGN + rekey: if a key or state of the PRNG is compromised, adding 1 bit or physical entropy does not help. Attacher can try both options (guess) and track the state. Therefore entropy should be mixed in large batches.

\subsection{Fortuna RNG}
\begin{definition}
	Fortuna RGN has a \emph{Generator} and \emph{Accumulator}.

	Generator is based on block cipher (AES) which encrypts a 128-bit counter (never overflows). After $2^{16}$ block we rekey and using the next 2 blocks as key.

	Trick: do not reset the counter to avoid short cycles.

	Accumulator: gets entropy from other sources and resets Generator periodically.
	Has $p_0, ..., p_{31}$ pods of entropy. External RNG is mixed to the pods in Round-Robin order: assign RNG to $p_i$, move to the next.
	Such process distributes RNG among the pods.

	After 100ms we mix pods into Generator. In the $j$-th mixing only selected pods are used, e.g. with index $2^j/i$.
\end{definition}

Assume that $\rho = $ rate of input entropy, measured in bits per 100ms.
If $\rho \geq 128 * 32 \Rightarrow p_0$ is enough for recovery. As at least 128 bit of entropy were put in $p_0$.
If $ \rho \geq 128 * 32/2^i \Rightarrow p_i$ is enough to recover.

To conclude, for any given rate of entropy Fortuna can recover after some time. Speed of recovery depends on entropy rate.

\subsection{Secure channel}
We use symmetric encryption to establish secure channel. Assume Alice and Bob share unique random key $k$.
Alice sends $\{ m_i \}_0^n$ sequence of messages, Bob receives a subsequence and he is still able to know which one he got.
On the other hand, attacker have no such info, except for length of the sequence.

We use
\begin{itemize}
	\item Random IV for every message (public).
	\item AES in CTR mode for encryption (alternatively ChaCha20).
	\item MAC after encrypt.
	\item Sequence numbers (inside MAC).
	\item Key derivation function which generates encryption and MAC keys for both directions.
		\[ h(k | id) \]
\end{itemize}
